\section{Equicontinuity}

\begin{definition}[Topological dynamical system]
	A \emph{topological dynamical system} (TDS) is a pair $(X, T)$, where
	\begin{itemize}
		\item $X$ is a non-empty, compact metric space,
		\item $T$ is a topological group,
		\item $\cdot : T \times  X \to X$ is a continuous group action, that is $\forall x \in X, \ s,t \in T$: 
		\begin{enumerate}
			\item $ e \cdot x = x$,
			\item $(st)\cdot x = s \cdot (t \cdot x)$.
		\end{enumerate}
	\end{itemize}
\end{definition}

\begin{proposition}[Subsystem]
	Let $(X, T)$ be a TDS and $A \subseteq X$ a closed, $T$-invariant set.
	Then $(A, T)$ is also a TDS with the inherited $T$-action called the \emph{subsystem} of $(X, T)$ induced by $A$.
\end{proposition}
\begin{proof}
	Since $A$ is a closed subset of a compact Hausdorff space it is a compact Hausdorff space. The invariance of $A$ implies that the $T$ action is well defined, continuity is clear. 
\end{proof}

\begin{definition}[Product system]
	Let $(X_\alpha, T)_{\alpha \in A}$ be an at most countable family of topological dynamical systems. Then the \emph{product system} $(X, T)$ of the family is defined by
	\begin{equation*}
		X := \prod_{\alpha \in A} X_\alpha, \quad t (x_\alpha)_{\alpha \in A} := (t x_\alpha)_{\alpha \in A}.
	\end{equation*}
\end{definition}

\begin{definition}[Morphism of TDS]
	Let $(X, T)$ and $(Y,T)$ be topological dynamical systems. A map $\pi : X \to Y$ is called a \emph{homomorphism} from $(X, T) \to (Y, T)$ if $\pi$ is continuous and $\forall x \in X, t \in T$:
	\begin{equation*}
		t \pi (x) = \pi(tx).
	\end{equation*}
	It is an isomorphism if $\pi$ is a homeomorphism.
\end{definition}

\begin{definition}[Factor of TDS]
	Let $(X, T)$ and $(Y, T)$ be topological dynamical systems and $\pi: X \to Y$ a homomorphism. Then $\pi$ is called a \emph{factor map} if it is surjective.
	If there exists a factor map $X \to Y$ then $Y$ is called a \emph{factor} of $X$. If $\pi$ is an isomorphism then $(X, T)$ and $(Y, T)$ are \emph{conjugates}.
\end{definition}

\begin{proposition}[Quotient system]
	Let $(X, T)$ be a TDS. Let $R \subset X \times X$ be a $T$-invariant (i.e. $\forall t \in T, \forall (x, x^\prime ) \in R : (tx, tx^\prime) \in R$) and closed equivalence relation (ICER). Then the \emph{quotient} $(X/R, T)$ with $T$-action defined by $t[x] := [tx]$ is a TDS.
\end{proposition}
\begin{proof}
	Let $\pi_R: X \to X/R$ be the projection, which is continuous. The quotient $X/R$ is a metrizable space, since $X$ is a metric space and $R$ is closed (not trivial). The compactness of $X/R$ follows from $X/R = \pi(X)$ and the compactness of $X$ together with the continuity of $\pi$. The $T$-invariance of $R$ implies the well-definedness of the $T$-action. To show that the $T$ action is continuous: Define the equivalence relation
	\begin{equation*}
		R^\prime := \{ ((t_1,x_1), (t_2,x_2)) \in (T \times X) \times (T \times X): (x_1, x_2) \in R, t_1 = t_2\}.
	\end{equation*}
	Then the $T$-action induces a continuous map $\cdot : (T \times X)/R^\prime \to (T\times X)/R^\prime$ by the universal property of the quotient topology. The continuous map $(t, x) \mapsto (t, [x])$ induces the continuous and bijective map $(T\times X)/R^\prime \ni [(t, x)] \mapsto (t, [x])$. This map is closed\footnote{this is because $x \mapsto [x]$ is closed and proper + thm 5.17.5 of \url{https://stacks.math.columbia.edu/tag/005M}}, which implies that it is a homeomorphism.
\end{proof}

\begin{proposition}[Connection between Factors and ICERs]
	\label{prop:factorsAndIcers}
	Let $(X, T)$ be a TDS. If $R \subset X \times X$ is an ICER, then the quotient map $\pi_R: X \to X/R$ defines a factor of map.
	
	If $(Y, T)$ is a factor of $(X, T)$ with factor map $\pi: X \to Y$, then
	\begin{equation*}
		R_\pi := \{(x, y) \in X \times X;\ \pi(x) = \pi(y)\}
	\end{equation*}
	defines an ICER and the systems $(Y, T)$ and $(X/R_\pi, T)$ are conjugated.
\end{proposition}
\begin{proof}
	The projection $\pi_R$ is continuous by definition of the quotient topology. For $x \in X$ and $t \in T$ we have
	\begin{align*}
		\pi_R(tx) = [tx] = t[x] = t\pi_R(x),
	\end{align*}
	so $\pi_R$ is indeed a factor map.
	
	It is clear that $R_\pi$ is closed, since $\pi$ is continuous. For $(x, y) \in R_\pi$ and $t \in T$ we have
	\begin{align*}
		\pi(tx) = t\pi(x) = t\pi(y) = \pi(ty),
 	\end{align*}
 	so $(tx, ty) \in R_\pi$, so $R_\pi$ is $T$-invariant. For $x \in X$ we have $\pi(x) = \pi(x)$ so $(x, x) \in R_\pi$, so $R_\pi$ is reflexive. Further for $(x, y) \in R_\pi$ we have $\pi(y) = \pi(x)$ so $(y, x) \in R_\pi$, which is therefore symmetric. Also for $x, y, z \in X$ with $(x, y) \in R_\pi$ and $(y, z) \i R_\pi$ we have $\pi(x) = \pi(y) = \pi(z)$, so $(x, z) \in R_\pi$ proving the transitivity of $R_\pi$. So we have shown that $R_\pi$ is an equivalence relation, hence an ICER.
 	
 	Define $\rho: X/R \to Y$ by $\rho([x]) = \pi(x)$. This is well defined, since $[x] = [y]$ implies $\pi(x) = \pi(y)$ by definition of $R_\pi$. For $U \subseteq Y$ open we have for all $x \in X$ that $\rho([x]) \in U$ if and only if $\pi(x) \in U$, so $[x] \in \rho^{-1}(U)$ if and only if $x \in \pi^{-1}(U)$. Thus $\pi_{R_\pi}^{-1}(\rho^{-1}(U)) = \pi^{-1}(U)$, which is open since $\pi$ is continuous. Thus by definition of the quotient topology $\rho^{-1}(U)$ is open, so $\rho$ is continuous. For $[x], [y] \in X/R_\pi$ with $\rho([x]) = \rho([y])$ we have $\pi(x) = \pi(y)$ so $[x] = [y]$, thus $\rho$ is injective. Also for $y \in Y$ there exists, since $\pi$ is surjective, a $x \in X$ such that $\pi(x) = y$. Therefore $\rho([x]) = \pi(x) = y$, proving that $\rho$ is surjective. Thus $\rho$ is bijective and since $X/R_\pi$ and $Y$ are compact Hausdorff $\rho$ is a homeomorphism. Lastly for $x \in X$ and $t \in T$
 	\begin{align*}
 		\rho(t[x]) = \rho([tx]) = \pi(tx) = t\pi(x) = t\rho([x])
 	\end{align*}
 	proves that $\rho$ is an isomorphism of TDSs.
\end{proof}

\begin{definition}[Equivalent and conjugated factors]
	Let $(X, T)$ be a TDS and factors $(Y, T)$ and $(Z, T)$ factors of it with factor maps $\pi: X \to Y$ and $\rho: X \to Z$.	Then the factors are called \emph(equivalent) if $R_\pi = R_\rho$ and \emph(conjugated) if $(Y, T)$ and $(Z, T)$ are conjugated.
	
	Similarly if $R$ and $S$ are two ICERs of $(X, T)$, they are called \emph{equivalent} if $R = S$ and \emph{conjugated} if $(X/R, T)$ and $(X/S, T)$ are conjugated.
\end{definition}

\begin{example}
	Let $(X, T)$ be a TDS. Let $\pi_i: X \times X \to X$ for $i = 1, 2$ be the projection on the $i$-th coordinate. Then $\pi_1$ and $\pi_2$ are factor maps of $(X \times X, T)$ to the factor $(X, T)$. They are clearly conjugates, as the factors are the same TDS. However it is easily verified that
	\begin{align*}
		R_{\pi_i} = \{((x_1, x_2), (y_1, y_2)) \in (X \times X)^2;\ x_i = y_i\}
	\end{align*}
	holds, so $R_{\pi_1} \neq R_{\pi_2}$, so the factors are not equivalent.
\end{example}

\begin{definition}[Equicontinuous TDS]
	Let $(X,T)$ be a TDS. Then $(X,T)$ is called \emph{equicontinuous}, if
	\begin{equation*}
		\forall \varepsilon > 0\ \exists \delta > 0: \forall x , y \in X, t \in T:  d(x, y) < \delta  \Rightarrow d(tx, ty) < \varepsilon.
	\end{equation*}
\end{definition}

\begin{example}[Dyadic odometer]
	Take $X = \{0, 1\}^\mathbb{N}$. We define a map $o: X \to X$ inductively by
	\begin{align*}
		o(a_1a_2a_3\dots) :=
		\begin{dcases}
			1a_2a_3\dots &;\ a_1 = 0\\
			0o(a_2a_3\dots) &;\ a_1 = 1
		\end{dcases}
	\end{align*}
	for all $a_1a_2a_3\dots \in X$. We will skip the proof that this gives rise to a well-defined map on $X$ and has the also inductively defined inverse
	\begin{align*}
		o^{-1}(a_1a_2a_3) :=
		\begin{dcases}
			1o^{-1}(a_2a_3\dots) &;\ a_1=0\\
			0a_2a_3\dots &:\ a_1=1
		\end{dcases}
	\end{align*}
	for all $a_1a_2a_3\dots \in X$.
	
	It is immediate from the definition that if $x, y \in X$ agree on the first $n$ entries, so do $o(x)$ and $o(y)$ as well as $o^{-1}(x)$ and $o^{-1}(y)$. Thus if we equip $X$ with the metric
	\begin{align*}
		d(a_1a_2\dots, b_1b_2\dots) := \min\{\frac{1}{n};\ a_n=b_n\}
	\end{align*}
	for $a_1a_2\dots, b_1b_2\dots \in X$ we have that $o$ and $o^{-1}$ are $1$-Lipschitz, which in turn implies that they must be isometries. Thus they are continuous and $(X, \mathbb{Z})$ is equicontinuous.
	
	We will skip the proof that every $x \in X$ has dense orbit under this $\mathbb{Z}$ action, so in particular $(X, \mathbb{Z})$ is minimal.
\end{example}

\begin{proposition}[Equicontinuity is independent on the choice of metric]
	Let $(X, T)$ be a TDS and $d, d'$ two metrics on $X$ that give rise to the same topology. Then $(X, T)$ is $d$-equicontinuous if and only if it is $d'$-equicontinuous.
\end{proposition}
\begin{proof}
	Fix $\varepsilon > 0$. The fact that $d$ and $d'$ give rise to the same topology, implies that $\mathrm{id}_X$ is $d$-$d'$ continuous. Since $X$ is compact it is even uniformly continuous, so there exists a $\tilde{\varepsilon} > 0$ such that
	\begin{align*}
		d(x, y) < \tilde{\varepsilon} \Rightarrow d'(x, y) < \varepsilon.
	\end{align*}
	Since $(X, T)$ is $d$-equicontinuous, there exists a $\tilde{\delta} > 0$ such that
	\begin{align*}
		d(x, y) < \tilde{\delta} \Rightarrow d(tx, ty) < \tilde{\varepsilon}
	\end{align*}
	holds for all $x, y \in X$ and $t \in T$. The identity map $\mathrm{id}_X$ is also $d'$-$d$ uniformly continuous, so there exists some $\delta > 0$ such that
	\begin{align*}
		d'(x, y) < \delta \Rightarrow d(x, y) < \tilde{\delta}
	\end{align*}
	holds for all $x, y \in X$. Thus we have
	\begin{align*}
		d'(x, y) < \delta \Rightarrow d(x, y) < \tilde{\delta} \Rightarrow d(tx, ty) < \tilde{\varepsilon} \Rightarrow d'(tx, ty) < \varepsilon
	\end{align*}
	for all $x, y \in X$ and $t \in T$. Hence $(X, T)$ is $d'$-equicontinuous.
\end{proof}

\begin{remark}
	The above proposition justifies, that we will often not give an explicit metric.
\end{remark}

\begin{proposition}[Isomorphism invariance of equicontinuity]
	\label{prop:equiContConjugationInvariance}
	Let $(X, T)$ and $(Y, T)$ be two isomorphic topological dynamical systems. If $(X, T)$ is equicontinuous, so is $(Y, T)$.
\end{proposition}
\begin{proof}
	Let $\pi: X \to Y$ an isomorphism and fix a metric $d_X$ on $X$. Then $d_Y: Y \times Y \to [0, \infty)$ defined by $d'(x, y):= d_X(\pi^{-1}(x), \pi^{-1}(y))$ for all $x, y \in Y$ is a metric on $Y$ turning $\pi$ into an isometry. For $r > 0$ and $x \in Y$ we have $B_r^{d_Y}(x) = \pi(B_r^{d_Y}(\pi^{-1}(x)))$, which is open, so the topology generated by $d_Y$ is finer than the inherent topology of $Y$. Also for $U \subseteq Y$ open and $x \in Y$, we have $\pi^{-1}(U)$ open, so there is some $r > 0$ such that $B_r^{d_X}(\pi^{-1}(x)) \subseteq \pi^{-1}(U)$, so $B_r^{d_Y}(x) = \pi(B_r^{d_Y}(\pi^{-1}(x))) \subseteq U$, which proves that $U$ is open in the topology generated by $d_Y$. Thus $d_Y$ generates the inherit topology of $Y$.
	
	For $\epsilon > 0$ there is some $\delta > 0$ such that $d_X(x, y) < \delta$ implies $d_X(tx, ty) < \epsilon$ for all $x, y \in X$ and $t \in T$. Then we have for $x, y \in Y$ with $d_Y(x, y) < \delta$ and $t \in T$ that $d_X(\pi^{-1}(x), \pi^{-1}(y)) = d_Y(x, y) < \delta$ so
	\begin{align*}
		d_Y(tx, ty) = d_X(\pi^{-1}(tx), \pi^{-1}(ty)) = d_X(t\pi^{-1}(x), t\pi^{-1}(y)) < \epsilon.
	\end{align*}
	This proves the equicontinuity of $(Y, T)$.
\end{proof}

\begin{proposition}[Products of equicontinuous systems]
	\label{prop:productsOfEquiContAreEquiCont}
	Let $(X_n, T)_{n \in \mathbb{N}}$ be a family of equicontinuous dynamical systems. Then the product $(\prod_{n \in \mathbb{N}} X_n, T)$ is also equicontinuous.
\end{proposition}
\begin{proof}
	Set $X:= \prod_{n \in \mathbb{N}} X_n$. Fix for each $n \in \mathbb{N}$ a metric $d_n$ of $X_n$ which gives rise to the topology of $X_n$. Since $X_n$ is compact we have $\operatorname{diam}(X_n) = \sup_{x, y \in X_n} d_n(x, y) < \infty$, so we can rescale $d_n$, such that $\operatorname{diam}(X_n) < 1$. Then
	\begin{align*}
		&d: X \times X \to [0, \infty) &d((x_n)_{n \in \mathbb{N}}, (y_n)_{n \in \mathbb{N}}) := \sum_{n \in \mathbb{N}} 2^{-n} d(x_n, y_n) &&\forall (x_n)_{n \in \mathbb{N}}, (y_n)_{n \in \mathbb{N}} \in X
	\end{align*}
	defines a metric on $X$, which gives rise to the product topology. Fix $\varepsilon > 0$. Then there is some $N \in \mathbb{N}$ such that $\sum_{n=N+1}^\infty 2^{-n} < \frac{\varepsilon}{2}$. For $n = 1, \dots, N$ there exists a $\delta_n > 0$ such that
	\begin{align*}
		d_n(x_n, y_n) < \delta_n \Rightarrow d_n(tx_n, ty_n) < \frac{\varepsilon}{2N}
	\end{align*}
	holds for all $x_n, y_n \in X_n$ and $t \in T$. Set $\delta := \min_{n=1, \dots, N} 2^{-n}\delta_n$. Fix $(x_n)_{n \in \mathbb{N}}, (y_n)_{n \in \mathbb{N}} \in X$ with $d((x_n)_{n, \in \mathbb{N}}, (y_n)_{n \in \mathbb{N}}) < \delta$ and $t \in T$. For $n = 1, \dots, N$ we have
	\begin{align*}
		d_n(x_n, y_n) = 2^n 2^{-n} d_n(x_n, y_n) \leq 2^n \sum_{m \in \mathbb{N}} 2^{-m} d_m(x_m, y_m) = 2^n d((x_m)_{m \in \mathbb{N}}, (y_m)_{m \in \mathbb{N}}) < 2^n \delta \leq 2^n 2^{-n} \delta_n = \delta_n,
	\end{align*}
	so $d_n(tx_n, ty_n) < \frac{\varepsilon}{2N}$. Thus we have
	\begin{align*}
		d(t(x_n)_{n \in \mathbb{N}}, t(y_n)_{n \in \mathbb{N}}) &= d((tx_n)_{n \in \mathbb{N}}, (ty_n)_{n \in \mathbb{N}})\\
		&= \sum_{n \in \mathbb{N}} 2^{-n} d_n(tx_n, ty_n)\\
		&= \sum_{n=1}^N 2^{-n} d_n(tx_n, ty_n) + \sum_{n=N+1}^\infty 2^{-n} d_n(tx_n, ty_n)\\
		&< \sum_{n=1}^N \frac{\varepsilon}{2N} + \sum_{n=N+1}^\infty 2^{-n}\\
		&< \frac{\varepsilon}{2} + \frac{\varepsilon}{2} &= \varepsilon,
	\end{align*}
	which proves that $(X, T)$ is equicontinuous.
\end{proof}
\begin{remark}
	The proposition above is clearly also true for finite products. If one extends the notion of equicontinuity from metric to uniform spaces it also holds for arbitrary products.
\end{remark}

\begin{proposition}[Subsystems of equicontinuous systems]
	\label{prop:SubsystemOfEquiContIsEquiCont}
	Let $(X, T)$ be an equicontinuous dynamical systems and $A \subseteq X$ a closed $T$-invariant set. Then $(A, T)$ is equicontinuous.
\end{proposition}
\begin{proof}
	This is clear.
\end{proof}
