\section{MEF}
\begin{definition}[Topological dynamical system]
  A \emph{topological dynamical system} (TDS) is a triplet $(X,T,\cdot)$, where
  \begin{itemize}
    \item $X$ is a non-empty, compact metric space,
    \item $T$ is a topological group,
    \item $\cdot : T \times  X \to X$ is a continuous group action, that is $\forall x \in X, \ s,t \in T$: 
    \begin{enumerate}
      \item $ e \cdot x = x$,
      \item $(s \cdot t)\cdot x = s \cdot (t \cdot x)$.
    \end{enumerate}
  \end{itemize}
\end{definition}
\begin{remark}
  The map $\cdot$ is usually supressed in the notation.
\end{remark}

\begin{definition}[Morphism of TDS]
  Let $(X,T), \ (Y,T)$ be topological dynamical systems.
  A map $\pi : X \to Y$ is called a \emph{homomorphism} from $(X,T) \to (Y,T)$ if $\pi$ is continuous and $\forall x \in X, t \in T$:
  \begin{equation*}
     t \pi (x) = \pi (t x).
  \end{equation*}
\end{definition}
\begin{definition}[Factor of TDS]
   Let $(X,T), \ (Y,T)$ be topological dynamical systems and $\pi: X \to Y$ a homomorphism.
   Then $\pi$ is called a \emph{factor map} if it is surjective.
   If there exists a factor map $X \to Y$ then $Y$ is called a \emph{factor} of $X$.
\end{definition}

\begin{proposition}[Quotient system]
  Let $(X,T)$ be a TDS. Let $R \subset X \times X$ be a $T$-invariant\footnote{that is $\forall t \in T, \forall (x,x^\prime ) \in R : (xt ,x^\prime t) \in R$} and closed equivalence relation (ICER for short).
  Let $\pi : X \to X/R$ be the quotient map.
  Endow $X/R$ with the quotient topology, that is $U \subset X/R$ open $: \Leftrightarrow \pi^{-1}(U)$ is open.
  Define the $T$-action by $t \cdot [x] := [t \cdot x]$.
  Then $(X/R, T)$ is a TDS and $\pi$ is a factor map.
\end{proposition}
\begin{proof}
  The quotient $X/R$ is a Hausdorff space, because $X$ is a compact Hausdorff space and $R$ is closed (not trivial).
 The compactness of $X/R$ follows from $X/R = \pi (X)$ and the compactness of $X$ together with the continuity of $\pi$. The $T$-invariance of $R$ implies the well-definedness of the $T$-action and that $\pi$ is a factor map.
 To show that the $T$ action is continuous: Define the equivalence relation
 \begin{equation*}
   R^\prime := \{ ((t_1,x_1),(t_2,x_2)) \in (T \times X) \times (T \times X): (x_1, x_2) \in R, t_1 = t_2\}.
 \end{equation*}
  Then the $T$-action induces a continuous map $\cdot : (T \times X)/R^\prime \to (T\times X)/R^\prime$ by the universal property of the quotient topology.
  The continuous map $(t,x) \mapsto (t,[x])$ induces the continuous and bijective map $(T\times X)/R^\prime \ni [(t,x)] \mapsto ( t,[x])$. 
  This map is closed\footnote{this is because $x \mapsto [x]$ is closed and proper + thm 5.17.5 of \url{https://stacks.math.columbia.edu/tag/005M}}, which implies that it is a homeomorphism.
\end{proof}


\begin{proposition}[ICER generated by factor]
  \label{prop:icer2factor}
  Let $(X,T)$ and $(Y,T)$ be TDS and $\pi : X \to Y$ a factor map.
  Then 
  \begin{equation*}
    R:= \{ (x_1,x_2)  \in X \times X :  \pi (x_1) = \pi(x_2) \}
  \end{equation*}
  is an ICER and the map $\tilde{\pi}: X / R \to Y$ induced by $\pi$ is an isomorphism $(X/R,T) \to (Y,T)$.
\end{proposition}

\begin{definition}[Product system]
  Let $(X_\alpha,T, \cdot_\alpha)_{\alpha \in A}$ be a family of topological dynamical systems.
  Then the \emph{product system} $(X, T, \cdot)$ of the family is defined by
  \begin{equation*}
    X := \prod_{\alpha \in A} X_\alpha, \quad (t \cdot x )_{\alpha} := t \cdot_\alpha x.
  \end{equation*}
\end{definition}

\begin{definition}[Equicontinuous TDS]
  Let $(X,T)$ be a TDS.
  Then $(X,T)$ is called \emph{equicontinuous}, if
  \begin{equation*}
  \forall \varepsilon \in (0, \infty) \exists \delta \in (0, \infty): \forall x , x^\prime \in X :  d(x, x^\prime) <\delta  \Rightarrow \forall t \in T: d(tx,tx^\prime) < \varepsilon.
  \end{equation*}
\end{definition}

\begin{proposition}[Stability of equicontinuity]
  products of eq. cont. systems are eq. cont., subsystems of eq. cont. systems are eq. cont., equicontinuity is isomorphism invariant
\end{proposition}


\begin{proposition}[Subsystem]
  Let $(X,T)$ be a TDS and $A \subset X$ a closed, $T$-invariant set.
  Then $(A,T)$ is also a TDS with the inherited $T$-action called the \emph{subsystem} of $(X,T)$ induced by $A$.
\end{proposition}
\begin{proof}
  Since $A$ is a closed subset of a compact Hausdorff space it is a compact Hausdorff space. The invariance of $A$ implies that the $T$ action is well defined, continuity is clear. 
\end{proof}

\begin{theorem}[Maximal equicontinuous factor]
  Let $(X,T)$ be a TDS.
  Then $(X,T)$ has an equicontinuous factor that is maximal in the sense that the ICER generated by it is the smallest with respect to set inclusion. This factor is called the \emph{maximal equicontinuous factor} (MEF) and it is unique up to isomorphism.
\end{theorem}
\begin{proof}
  Existence:
  Let
  \begin{equation*}
    \mathcal{R} := \{ R \subset X \times X : R \ \text{ICER with} \ X/R \ \text{equicontinuous}\}.
  \end{equation*}
  Then $ X \times X \in \mathcal{R}$ and so the set is non empty.
  If $\varphi : (X,T)\to (Y,T)$ is a factor with $(Y,T)$ equicontinuous and $R$ the Relation generated by $\varphi$, then $X/R$ is also equicontinuous, since $(X/R,T)$ is isomorphic to $(Y,T)$.
  Now let 
  \begin{equation*}
    R:= \cap \mathcal{R}.
  \end{equation*}
  Then $R$ is an ICER (because intersection of ICERs are ICERs).
  \textbf{Hier fehlt noch das Argument für die Metrisierbarkeit.}
  The product system $(\prod_{R^\prime \in \mathcal{R}} X/R^\prime,T)$ is equicontinuous, since products of equicontinuous systems are equicontinuous.
Consider the morphism $f:= X \ni x  \mapsto ([x]_{R^\prime})_{R^\prime \in \mathcal{R}}$.
  Let $A:= f(X)$, then $A$ is $T$-invariant being the image of a morphism.
  Furthermore $A$ is compact by continuity of $f$ and compactness of $X$.
  Since $A$ is a compact subset of a Hausdorff space it follows that $A$ is closed.
  Therefore we can consider the subsystem $(A,T)$ which is an equicontinuous system (subsystems of equicontinuous systems are equicontinuous).
  Consider the map $\pi: X \to  A , \  x  \mapsto ([x]_{R^\prime})_{R^\prime \in \mathcal{R}}$.
  Then $\pi$ is a factor and the ICER generated by $\pi$ is equal to $R$.
  Uniqueness:
  Any maximal equicontinuous factor generates the ICER $R$ and any two factors that generate $R$ are isomorphic to $(X/R,T)$ (proposition \ref{prop:icer2factor}).
\end{proof}
