\section{Maximal equicontinuous factor}
\subsection{Definition, $\exists!$, universal property}
\begin{frame}
  Let $(X,T)$ be a TDS.
  \begin{definition}[maximal equicontinuous factor (MEF)]
    A factor $\pi : (X,T) \to (X_{\text{MEF}},T)$ is a \textbf{MEF} of $(X,T)$ if and only if
  \begin{itemize}
    \item $(X_{\text{MEF}},T)$ is equicontinuous,
    \item $\pi$ is maximal: $\forall \varphi : (X,T) \to (Y,T)$ factor s.t. $(Y,T)$ equicontinuous: $R_\pi \subseteq R_\varphi$.
  \end{itemize}
\end{definition}
\pause
  \begin{alertblock}{Theorem (existence and uniqueness of the MEF)}
  $(X,T)$ has a MEF that is unique up to equivalence.
  \end{alertblock}
  \pause
\begin{example}[MEF of equicontinuous TDS is original TDS]
  $(X,T)$ equicontinuous, then $I: (X,T) \to (X,T)$ is the MEF of $(X,T)$.
\end{example}

\end{frame}
\begin{frame}[fragile]
\begin{proposition}[universal property of the MEF]
  Let $\pi : (X, T) \to  (X_{\text{MEF}},T)$ be the MEF.
  \begin{equation*}
    \begin{split}
      &\forall \, \varphi : (X,T) \to (Y,T) \ \text{factor with} \ (Y,T) \ \text{equicontinuous}: \\
      & \exists! \text{factor} \ \tilde{\varphi}: (X_{\text{MEF}},T) \to (Y,T): \varphi = \tilde{\varphi} \circ \pi
    \end{split}
      \end{equation*}
  \end{proposition}
  \begin{center}
    \begin{tikzcd}[sep=large]
      (X,T) \arrow[d, "\pi"'] \arrow[r, "\varphi"]  \arrow[dr, phantom, "\circlearrowleft", near start] 
      & (Y,T) \ \text{equic.}\\
      (X_{\text{MEF}},T)  \arrow[ur,"\exists ! \tilde{\varphi}"'] & \phantom{a} \\
    \end{tikzcd}
  \end{center}
\end{frame}
\subsection{Eigenfunction characterisation and $C^*$-algebra approach}
\begin{frame}
  $(X,T)$ TDS with $T$ abelian, $T^\sharp$ ($T^*$) the set of all (continuous) characters of $T$.
\begin{definition}[Koopman representation]
  \begin{equation*}
    \begin{split}
      &U : T \longrightarrow \mathscr{L}(C(X)) \\
      &(U(t) f)(x) \coloneq f(tx)
    \end{split}
  \end{equation*}
   \end{definition}
   \begin{remark}
    $T$-action continuous on $T \times X$ $\Rightarrow$ $U$ strongly continuous $T$-representation

    \hfill(for proof see \cite{OpThErg}, Theorem 4.17)
   \end{remark}
   \pause
   \begin{definition}[eigenvalue, eigenfunction]
   \begin{equation*}
    \begin{split}
      &0 \neq f \in C(X) \ \text{\emph{eigenfunction} of} \ U \ \text{to \emph{eigenvalue}} \ \chi \in T^*   \\
 &:\Leftrightarrow f \in \ker (U-\chi)\coloneq \bigcap_{t \in T} \ker (U(t)- \chi (t)I)
    \end{split}
      \end{equation*}
   \end{definition}
\end{frame}
\begin{frame}[fragile]
  \begin{definition}[discrete spectrum]
  \begin{equation*}
    (X,T) \ \text{has (TDS) \textbf{discrete spectrum}}:\Leftrightarrow C(X) = \overline{\lin} \bigcup_{\chi \in T^*} \ker (U- \chi)
  \end{equation*}
\end{definition}
\pause 
  \begin{alertblock}{Theorem (equicontinuity $\Leftrightarrow$ discrete spectrum)}%
  \begin{equation*}
    (X,T) \ \text{is equicontinuous} \Leftrightarrow (X,T) \ \text{has discrete spectrum}
  \end{equation*}

    \hfill(for proof see \cite{HK2023}, Theorem 2.11)
\end{alertblock}
\pause
  \begin{alertblock}{Theorem (MEF eigenfunction characterisation)}
  \label{thm:MEF_EFchar}
  Let $\pi : (X,T) \to (X_{\text{MEF}},T)$ be the MEF.
  Then $\forall x_1, x_2 \in X$:
  \begin{equation*}
  \pi (x_1) = \pi (x_2) \Leftrightarrow 
    \forall f \in \bigcup_{\chi \in T^*} \ker (U- \chi) : f(x_1) = f(x_2)
  \end{equation*}
\end{alertblock}
\end{frame}
\begin{frame}{Intermezzo: Duality of categories!}
  \begin{itemize}
    \item $\mathbf{C}\coloneq$ category of commutative unital $C^*$-algebras,
    \item $\mathbf{D} \coloneq$ category of compact Hausdorff spaces,
          \item \pause
contravariant Gelfand functor $G: \mathbf{C} \to \mathbf{D}$,
      \begin{equation*}
        \begin{split}
          & G(\mathscr{A}) \coloneq \{ \varphi : \mathscr{A} \to \mathbb{C} \ | \ \varphi \ \text{morphism}\} \ \text{with the weak-* topology}, \\
          & G( \Psi : \mathscr{A} \to \mathscr{B}) : G(\mathscr{B} ) \to G(\mathscr{A}), \ \varphi \mapsto \varphi \circ \Psi,
        \end{split}
    \end{equation*}
  \item \pause contravariant continuous function functor $C: \mathbf{D} \to \mathbf{C}$, $X \mapsto C(X)$, 
    \begin{equation*}
       C(f:X \to Y) : C(Y) \to C(X), \ g \mapsto g \circ f,
    \end{equation*}
  \item \pause  evaluation map $\ev : I \Rightarrow GC$, $\ev (x) f\coloneq f(x)$ is natural isomorphism,
  \item Gelfand transform $\widehat{\phantom{a}} : I \Rightarrow CG$, $\widehat{a}(\varphi)\coloneq\varphi (a)$ is natural isomorphism.
      \end{itemize}
      \end{frame}
\begin{frame}[fragile]{MEF via $C^*$-algebras}
  \begin{columns}[t]
    \begin{column}{0.5\textwidth}
     \begin{itemize}
       \item $\mathscr{A}\coloneq \clin \bigcup_{\chi \in T^*} \ker (U- \chi) \subseteq C(X)$,
    \item  $\mathscr{A}$ is smallest unital $C^*$-subalgebra containing all eigenfunctions of $U$,
   \item \pause
      $\mathscr{A}$ $U$-invariant $\rightarrow T$-action on $G(\mathscr{A})$:
      \begin{equation*}
      t\cdot \varphi \coloneq G(U(t))\varphi,
      \end{equation*}
    \item \pause $\iota : \mathscr{A} \to C(X)$ natural injection,
    \item $G(\iota)  : G(C(X)) \to G(\mathscr{A})$,
    \item \alert{MEF of $(X,T)$ is}
      \begin{equation*}
        \begin{split}
          &\alert{\pi : (X,T) \to (G(\mathscr{A}),T),}\\ 
          &\alert{\pi \coloneq G(\iota) \circ \ev,}
        \end{split}
      \end{equation*}
        \end{itemize}
    \end{column}
    \begin{column}{0.5 \textwidth}
 \begin{itemize}
 \item \pause $\widehat{\phantom{a}}: I \Rightarrow CG$ natural:
    \begin{tikzcd}[sep=large]
      \mathscr{A} \arrow[d, "\iota"'] \arrow[r, "\widehat{\phantom{a}}"]  \arrow[dr, phantom, "\circlearrowleft"] 
      & C(G(\mathscr{A})) \arrow[d,"C(G(\iota))"] \\
      C(X)  \arrow[r,"\widehat{\phantom{a}}"] & C(G(C(X)))
    \end{tikzcd}
   \item \pause
          $G(\iota)$ surjective $\Leftrightarrow$ $C (G( \iota ))$ injective $\Leftrightarrow$ $\iota$ injective,
    \item \pause let $x_1, x_2 \in X$:
      \begin{equation*}
        \begin{split}
          &\pi (x_1) = \pi (x_2) \Leftrightarrow \forall f \in \mathscr{A} : f(x_1) = f(x_2) \\
          &\Leftrightarrow \forall f \in \bigcup_{\chi \in T^*} \ker (U- \chi)  : f(x_1) = f(x_2).
        \end{split}
    \end{equation*}
        \end{itemize}
    \end{column}
  \end{columns}
  \end{frame}
\subsection{Relationship to measure preserving system concepts}
\begin{frame}{MEF vs. Kronecker subsystem}
  \onslide<1>$T$ still abelian
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \onslide<1->$(X,T)$ TDS
      \begin{itemize}
        \item \onslide<2-> Koopman rep. $U : T \to \mathscr{L}(C(X))$,
        \item \onslide<4->$(X,T)$ has discrete spectrum $\Leftrightarrow \clin \bigcup_{\chi \in T^*} \ker (U- \chi) = C(X)$,
        \item \onslide<6-> MEF: associated to the $U$-invariant $C^*$-subalgebra
          \begin{equation*}
            \clin \bigcup_{\chi \in T^*} \ker (U- \chi)  \subseteq C(X).
          \end{equation*}
      \end{itemize}
    \end{column}
\begin{column}{0.5\textwidth}
  \onslide<1->
  $(Y,T)$ measure preserving system (MPS)
  \begin{itemize}
    \item \onslide<3-> Koopman rep. $V: T \to \mathscr{L}(L^2(Y))$,
    \item \onslide<5->$(Y,T)$ has discrete spectrum $\Leftrightarrow \clin \bigcup_{\chi \in T^\sharp} \ker (V- \chi) = L^2(Y)$,
    \item \onslide<7-> Kronecker subsystem: associated to the $V$-invariant Markov sublattice 
      \begin{equation*}
        \clin \bigcup_{\chi \in T^\sharp} \ker (V- \chi)  \subseteq L^2(Y).
      \end{equation*}
  \end{itemize}
    \end{column}
  \end{columns}
  \onslide<8->
  \centering
  \phantom{a}\\
  \textbf{MEF is the TDS analogue of the Kronecker subsystem!}
\end{frame}
\begin{comment}
\begin{frame}{TDS with invariant measure as MPS}
Let $(X,T)$ a TDS, $T$ abelian, $\mu$ a $T$-invariant probability measure on $X$.
\begin{itemize}
  \item $(X,T,\mu)$ is a concrete measure preserving system,
      \item \pause MPS Koopman representation
\begin{equation*}
  \begin{split}
    &V : T \to \mathscr{L}(L^2(X)),\\
    & (V(t) f)(x) \coloneq f (t x).
  \end{split}
\end{equation*}
\end{itemize}
\pause
  \begin{proposition}[TDS discrete spectrum $\Rightarrow$ MPS discrete spectrum]
  \begin{equation*}
    (X,T) \ \text{has TDS discrete spectrum} \Rightarrow (X,T,\mu) \ \text{has MPS discrete spectrum}
  \end{equation*}
  \end{proposition}
\end{frame}
\begin{frame}[fragile]{Relationship of MEF and Kronecker subsystem}
  Let $\pi : (X,T) \to (X_{\text{MEF}},T)$ be the MEF of $(X,T)$.
  \begin{itemize}
    \item $(X_{\text{MEF}},T, \nu)$ with $\nu(A)\coloneq  \mu(\pi^{-1}(A))$ is concrete MPS,
    \item $(X_{\text{MEF}},T, \nu)$ has MPS discrete spectrum,
    \item \pause let $\mathcal{X}, \mathcal{X}_m$ the MPS associated to $(X,T,\mu), (X_{\text{MEF}},T,\nu)$,
  \item let $\pi^* : \mathcal{X}_m \to \mathcal{X}$ the extension induced by $\pi$,
  \item \pause let $K: \mathcal{K} \to \mathcal{X}$ the Kronecker subsystem of $\mathcal{X}$,
  \item Kronecker subsystem maximality: $\exists$ extension $J: \mathcal{X}_m \to \mathcal{K}$:
\end{itemize}
  \begin{center}
    \begin{tikzcd}[sep=large]
      \phantom{a}       & \mathcal{K} \arrow[d,"K"]  \\
      \mathcal{X}_m \arrow[r,"\pi^*"'] \arrow[ur,"J"]  &  \arrow[ul, phantom, "\circlearrowright", near start] 
\mathcal{X}
    \end{tikzcd}\\
\pause
  \textbf{MEF factors through Kronecker subsystem!}
  \end{center}
  \end{frame}
\end{comment}
